### 2011-04-10 Thesis/tese.tex
Thesis upload

%%--------------------------------------------------
%%
%% Dissertação de Mestrado
%% Autor: Bruno Luna - Data(Início): 29/11/2010 
%%
%%---------------------------------------------------
\documentclass[pt,oneside,onehalfspacing]{risethesis}

\usepackage{natbib}
\usepackage{babel}
\usepackage{supertabular}
\usepackage{fancybox}
\usepackage{acronym}
%% Change the following pdf author attribute name to your name.
\usepackage[linkcolor=black,citecolor=black,urlcolor=black,colorlinks,pdfpagelabels,pdftitle={brunoluna-msc},pdfauthor={Bruno Gustavo Borges Luna}]{hyperref}

\address{RECIFE}

\universitypt{Universidade Federal de Pernambuco}
\universityen{Federal University of Pernambuco}

\departmentpt{Centro de Tecnologia e Geociências}
\departmenten{Centre for Technology and Geosciences}

\programpt{Pós-graduação em Engenharia Mecânica}
\programen{Graduate in Mechanical Engineering}

\majorfieldpt{Engenharia Mecânica}
\majorfielden{Mechanical Engineering}

\title{Aceleração de Convergência via Multigrid de Simulador de Escoamentos em Meios Porosos}
\date{JULHO/2011}

\author{Bruno Gustavo Borges Luna}
\adviser{Paulo Roberto Maciel Lyra}
\coadviser{Ramiro Brito Willmersdorf}

\begin{document}

\frontmatter
\frontpage
\presentationpage

\begin{dedicatory}
Dedico esta dissertação à \ldots
\end{dedicatory}

\acknowledgements
Gostaria de agradecer \ldots

\resumo
\input{chapters/extras/resumo}

\abstract
\input{chapters/extras/abstract}

\tableofcontents
\listoffigures
\listoftables
% Acronyms
%\input{chapters/extras/acronyms}
\mainmatter

%% Here you include your chapters
% \include{chapters/introduction/main}

% CHAPTER 01 - INTRODUCTION
\include{chapters/ch01/introduction}

% CHAPTER 02 - MATHEMATICAL FORMULATION
\include{chapters/ch02/formulacaoMatematica}

% CHAPTER 03 - NUMERICAL FORMULATION
\include{chapters/ch03/formulacaoNumerica}

% CHAPTER 04 - IMPLEMENTATION
\include{chapters/ch04/implementacao}

% CHAPTER 05 - RESULTS
\include{chapters/ch05/resultados}

\bibliographystyle{natbib}
\addcontentsline{toc}{chapter}{Referências Bibliográficas}
%% REFERENCES
\bibliography{chapters/extras/references}

% Appendix (Uncomment for using)
%\clearpage
%\addappheadtotoc
%\appendix
%\appendixpage
%% Here you include your appendix files
%\include{chapters/appendix/a01}

\end{document}