### 2011-04-10 Thesis/chapters/extras/abstract.tex
New structure

Multigrid methods allow a faster convergence of linear and non-linear equations systems solvers. However, they require a sequence of progressive coarser meshes whose generation can become a issue if not adequately addressed. Traditionally this sequence has been made setting different levels of refinement, manually or automatically, starting with the coarsest mesh. Nevertheless this approach often leads to difficulties in representing the geometrical details in the first mesh. It can also demand successive searches in order to perform interpolations between levels, introducing extra approximation errors. An alternative to overcome these problems is to start the mesh sequence generation from the finest level and gradually coarse it using an appropriate technique. This work presents the implementation of a control volume agglomeration algorithm for the fully automatic creation of consistent unstructured meshes applied to the numerical solution of porous media flows using a multigrid accelerated edge-based finite volume formulation. The methods described here are general enough to handle with three-dimensional and heterogeneous problems. Examples are shown and results discussed for two-dimensional model problems in homogenous domains.

\begin{keywords}
Multigrid, Finite Volume, Porous Media Flow
\end{keywords}