### 2011-04-19 Thesis/chapters/extras/abstract.tex
Adicionada funcao (a principio comentada, só descomenta quando precisar..., tem que ajeitar isto!) para plotar e imprimir na tela os autovalores da matriz A em Ax=b, os quais tem uma relacao direta com os autovalores da matriz de iteracao e portanto com o raio espectral. Tambem foram escritas as primeiras versoes dos resumos tanto em ingles quanto em portugues

The simulation of multiphase flows in porous media impose many numerical challenges due to a series of factors as the high anisotropic and heterogeneous media handled in this type of analysis, the coupled elliptic-hyperbolic mathematical nature of the associated Partial Differential Equations (PDEs),  among others. One of the main difficulties faced by numerical simulations of PDEs is the solution of large systems of linear equations. Multigrid methods allow a faster convergence of linear and non-linear equations systems solvers. However, they require a sequence of progressive coarser meshes whose generation can become a issue if not adequately addressed. Traditionally this sequence has been made setting different levels of refinement, manually or automatically, starting with the coarsest mesh. Nevertheless this approach often leads to difficulties in representing the geometrical details in the first mesh. It can also demand successive searches in order to perform interpolations between levels, introducing extra approximation errors. An alternative to overcome these problems is to generate all the mesh hierarchy in a full algebric manner using the so-called Algebraic Multigrid(AMG) approach. This work presents the implementation of a software created using the FEniCs tool for automated low-level code generation and PyAMG for the fully automatic creation of consistent multilevel matrix structures to be applied to the numerical solution of porous media flows using the Finite Element Method (FEM). The methods described here are general enough to handle with three-dimensional, heterogeneous and anisotropic problems. Examples are shown and results discussed for two-dimensional problems in homogenous and heterougenous domains with iso- and anisotropic tensors.

\begin{keywords}
Multigrid, Oil Reservoir, Porous Media Flow
\end{keywords}