### 2011-12-05 Thesis/tese.tex

%%--------------------------------------------------
%%
%% Dissertação de Mestrado
%% Autor: Bruno Luna - Data(Início): 29/11/2010 
%%
%%---------------------------------------------------
\documentclass[pt,oneside,onehalfspacing]{risethesis}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% PACKAGES AND CUSTOMIZATIONS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{natbib}
\usepackage{babel}
\usepackage{supertabular}
\usepackage{subfigure}
\usepackage{fancybox}
\usepackage{acronym}
% Commands for list of symbols
\usepackage[intoc,portuguese]{nomencl}
\makenomenclature
\renewcommand{\nomname}{Lista de Símbolos}
% Commands for code listings
\usepackage{codehighlight}
\usepackage{minted}
\usemintedstyle{tango}
\newminted{xml}{bgcolor=yellow!10, frame=single, rulecolor=\color{black!50}, label="XML"}
\newminted{python}{bgcolor=yellow!10, frame=single, rulecolor=\color{black!50}, label="Python"}
\renewcommand\listingscaption{Listagem}
%% Change the following pdf author attribute name to your name.
\usepackage[linkcolor=black,citecolor=black,urlcolor=black,colorlinks,pdfpagelabels,pdftitle={brunoluna-msc},pdfauthor={Bruno Gustavo Borges Luna}]{hyperref}
% Command added to present problem in 'theorem form'
\newtheorem{problem}{Problema}
% Definition for tensors format
\def\mathbi#1{\textbf{\em #1}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% FRONT MATTER
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\address{Recife}

\universitypt{Universidade Federal de Pernambuco}
\universityen{Federal University of Pernambuco}

\departmentpt{Centro de Tecnologia e Geociências}
\departmenten{Centre for Technology and Geosciences}

\programpt{Pós-graduação em Engenharia Mecânica}
\programen{Graduate in Mechanical Engineering}

\majorfieldpt{Engenharia Mecânica}
\majorfielden{Mechanical Engineering}

\title{Modelagem Automática de Escoamentos em Meios Porosos via Método dos Elementos Finitos}
\date{Janeiro/2012}

\author{Bruno Gustavo Borges Luna}
\adviser{Paulo Roberto Maciel Lyra}
\coadviser{Ramiro Brito Willmersdorf}

\begin{document}

\frontmatter
\frontpage
\presentationpage

\begin{dedicatory}
Dedico à \ldots
\end{dedicatory}

\acknowledgements
Gostaria de agradecer \ldots

\resumo
\input{chapters/extras/resumo}

\abstract
\input{chapters/extras/abstract}

\tableofcontents
\listoffigures
\listoftables
% ACRONYMS
\input{chapters/extras/acronyms}
\printnomenclature
\mainmatter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% BEGIN CHAPTERS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% CHAPTER 01 - INTRODUCTION
\include{chapters/ch01/introduction}

% CHAPTER 02 - MATHEMATICAL FORMULATION
\include{chapters/ch02/formulacaoMatematica}

% CHAPTER 03 - NUMERICAL FORMULATION
\include{chapters/ch03/formulacaoNumerica}

% CHAPTER 04 - IMPLEMENTATION
\include{chapters/ch04/implementacao}

% CHAPTER 05 - RESULTS
\include{chapters/ch05/resultados}

% CHAPTER 06 - CONCLUSION
\include{chapters/ch06/conclusoes}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% BACK MATTER
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{natbib}
\addcontentsline{toc}{chapter}{Referências Bibliográficas}
% REFERENCES
\bibliography{chapters/extras/references}

% APPENDIX
\clearpage
\appendix
% A01 - USER GUIDE
\include{chapters/appendix/a01}

% A02 - SELECTED SECTIONS OF CODE
\include{chapters/appendix/a02}

\end{document}