### 2012-02-02 Thesis/chapters/ch06/conclusoes.tex
Resumo+ch03(form.numerica)+resultados hiperbolico

\chapter{Conclusões e Trabalhos Futuros}
\label{ch:conclusoes}

\section{Conclusões}
\label{sc:conclusoes}
\emph{DESCRIÇÃO PENDENTE}

\section{Trabalhos Futuros}
\label{sc:trabFuturos}
\emph{DESCRIÇÃO PENDENTE}