### 2012-05-19 Thesis/packcustom.tex

\usepackage{natbib}
\usepackage{babel}
\usepackage{supertabular}
\usepackage{subfigure}
\usepackage{fancybox}
\usepackage{acronym}
\usepackage{url}
% Command to draw lines
\usepackage{tikz}
% Commands for list of symbols
\usepackage[intoc,portuguese]{nomencl}
\makenomenclature
\renewcommand{\nomname}{Lista de Símbolos}
% Commands for code listings
\usepackage{codehighlight}
\usepackage{minted}
\usemintedstyle{tango}
\newminted{xml}{bgcolor=yellow!10, frame=single, rulecolor=\color{black!50}, label="XML"}
\newminted{python}{bgcolor=yellow!10, frame=single, rulecolor=\color{black!50}, label="Python"}
\renewcommand\listingscaption{Listagem}
%% Change the following pdf author attribute name to your name.
\usepackage{float}
\usepackage[linkcolor=black,citecolor=black,urlcolor=black,colorlinks,pdfpagelabels,pdftitle={brunoluna-msc},pdfauthor={Bruno Gustavo Borges Luna}]{hyperref}
% Command added to present problem in 'theorem form'
\newtheorem{problem}{Problema}
% Definition for tensors format
\def\mathbi#1{\textbf{\em #1}}
% Command for algorithms in pseudocode
\usepackage{algorithm}
\usepackage{algorithmic}